---
title: "CMSC320 Final Project"
---

```{r setup, include=FALSE}
library("tidyverse")
```

# Introduction

# Loading and Pre-processing Data

Data can be easily loaded using R's `read.csv` method. For example, here we can load and display
`books.csv` to get a broad picture of one of the datasets we will be working
with. We can see that `books.csv` organizes its data in a tabulated manner with each book as a row
and the attributes of a book as columns.

```{r}
books_df <- read.csv("./goodbooks-10k/books.csv")
head(books_df)
```

Let's load the rest of our data as well.

```{r}
ratings_df <- read.csv("./goodbooks-10k/ratings.csv")
tags_df <- read.csv("./goodbooks-10k/tags.csv")
book_tags_df <- read.csv("./goodbooks-10k/book_tags.csv")
```

One of the things we are interested in is the genre of certain books.
We can narrow down the genres we are interested in with pipeline that
filters for certain genre names.

```{r}
tags_df <- tags_df %>%
  filter(tag_name %in% c(
    "fantasy",
    "mystery",
    "horror",
    "realistic-fiction",
    "romance",
    "science-fiction",
    "thriller",
    "young-adult"
  ))
tags_df
```
However, this information is not very helpful by itself. We
want to be able to use it in conjunction with the other data
we have. One way to accomplish this is by joining data across
different datasets through a common attribute. Here, we construct a pipeline
to link data from `books_df` and `book_tags_df` through the shared
attribute, `goodreads_book_id`, and then `book_tags_df` and `tags_df`
through `tag_id`. We can see that `tag_name` does indeed describe
the genre of the books it's associated with.

```{r}
books_df %>%
  inner_join(book_tags_df, by="goodreads_book_id") %>%
  inner_join(tags_df, by="tag_id") %>%
  select("title", "tag_name") %>%
  head(15)
```

# Data Analysis

# Predictions

# Conclusion

# Acknowledgements